---
title: "The Big-mac Index"
subtitle: <font size='6'>Purchasing Power Parity </font>
author: <font size='8'>The Code Ninjas <br> Tuilelaith, Dara, Kay Thari, Anya </font>
institute: <font size='10'>University of Edinburgh</font>
date: <font size = '6'>`r Sys.Date()`</font>
output:
  revealjs::revealjs_presentation:
    self_contained: false
editor_options: 
  chunk_output_type: console
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(rmarkdown)
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(revealjs)
library(gganimate)
library(gifski)
library(grid)
library(gridExtra)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
suppressWarnings(expr)
```

```{r load-data, include=FALSE}
big_mac_full_index = read.csv(file = "/cloud/project/data/big-mac-data-2021-07.1/output-data/big-mac-full-index.csv")
```

## <span style="color: turquoise;"> A statement of the overall goal / research question </span>  {data-background-image="img/Global-Economy-4.jpg"}
<span style="color: yellow;"> <font size='7'> Our research question is “What is Purchasing Power Parity (PPP) and how has it changed over the years across countries?” </font> </span>

# <span style="color: yellowgreen;"> What is Purchasing Power Parity? </span> {data-background-image="img/World.jpg"}

<span style="color: gold;"> <font size='6'> PPP stands for Purchasing Power Parity.</font></span>

<span style="color: gold;"> <font size='6'> It is a measurement used in economics to compare the purchasing power of different currencies in common units. </font></span>

<span style="color: gold;"> <font size='6'> In theory, PPP should be equal to the exchange rate of these two countries, but in practice this is not the case.</font></span>

# <span style="color: royalblue;"> The Big-mac Index Data</span> {data-background-image="img/Burger2.jpg"}

<font size='6'> The dataset consists of 1,520 observations (rows) and 19 variables (columns).</font>

 <font size='6'>Each row represents a country at a specific time (year-month-date) while the columns represent crucial information related to the countries.</font>

<font size='6'>The data is accountable for the years 2000-2021. </font>

<font size='6'>The Big Mac prices are from McDonald’s, and the exchange rates are from Thomson Reuters. The GDP per person data is from the IMF World Economic Outlook reports.</font>

---

```{r tibble}
dim(big_mac_full_index)

ls(big_mac_full_index)

```

<<<<<<< HEAD
# <span style="color: seagreen;"> Why Big Macs?</span> {data-background-image="img/Yellow gradient.jpg"}
=======

# <span style="color: yellow;"> Why Big Macs?</span> {data-background-image="img/Brown-Burger.jpg"}

<span style="color: snow;"> <font size='6'> The Big Mac Index is an accurate way of looking at PPP in the real world. </font></span>

<span style="color: snow;"> <font size='6'> A Big Mac is going to be the same and hold the same intrinsic value no matter what country it is in, and so it is a good product to compare the currencies across countries.</font></span>

# <span style="color: gold;"> Analysis of PPP</span> {data-background-image="img/Solar.jpg"}

<span style="color: turquoise;"><font size='6'>One of the common assumptions surrounding PPP is that a lower income country will on average, have lower prices compared to a high income country.</font></span>

<span style="color: turquoise;"><font size='6'>This could be due to lower labour costs and an overall inability to afford more costly goods in a low income economy. This will mean that a person in a low income economy will be able to buy less in a high income economy than they could before.</font></span>

<span style="color: turquoise;"><font size='6'>This would therefore give us the impression that PPP is lower in low income economies than in high income economies.</font></span>


# Plots - PPP by GDP per capita
::: columns
:::: column

- <font size='5'> The assumption that PPP is higher in high income countries is supported by this graph. Raw Index shows how overvalued/undervalued a countries currency is compared to the big mac price in each country.</font>

- <font size='5'> Effectively a country with a highly overvalued currency will have a high PPP and a country with a highly undervalued currency will have a low PPP.</font>


::::

:::: column
```{r plot-gdp/usdraw, echo = FALSE, warning=FALSE, out.height="100%", fig.height=5, message=FALSE}

bigmac2011 <- big_mac_full_index %>%
  filter(date >= as.Date("2011-01-01"), date <= as.Date("2011-12-31"))

ggplot(data = bigmac2011, mapping = aes(x = GDP_dollar, y = USD_raw )) +
  geom_point() +
  theme_minimal() +
  geom_smooth(method = "lm", se = FALSE, colour = "grey", linetype = "dashed",formula = y ~ x) +
  labs(
    x = "GDP per person (dollars)",
    y = "Raw Index (dollars)",
    title = "",
    subtitle = "For 2011",
    fill = "Countries"
  ) +
  theme(
    legend.position = "bottom"
  )



##line <- linear_reg() %>%
  ##set_engine("lm") %>%
  ##fit(USD_raw ~ GDP_dollar, data = bigmac2011) 


```
::::
:::

---

```{r plot-gdp/usdrawYRS, echo=FALSE, warning=FALSE, message=FALSE}


big_mac_full_index$year = str_sub(big_mac_full_index$date,1,4)

bigmac <- big_mac_full_index %>%
  filter(year %in% c("2012","2014","2016","2018","2020"))

bigmac %>%
  group_by(year) %>%
  ggplot(data = bigmac, mapping = aes(x = GDP_dollar, y = USD_raw, colour = name)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, colour = "grey", linetype = "dashed",formula = y ~ x) +
  labs(
    x = "GDP per person (dollars)",
    y = "Raw Index (dollars)",
    title = "",
    subtitle = "Through the decade"
  ) +
  theme(
    legend.position = "none",
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "lightgrey"),
    plot.title = element_text(face = "bold")
  )  +
  facet_grid(year ~ .)

```


## Turkey
::: columns
:::: column
- <font size='5'>High rates of GDP growth till 2007 with greater investor confidence.</font>

- <font size='5'>Post 2008, industry declined and led to higher levels of inflation and a reduction in imports, all of this led to the Lira fall in value up till 2021</font>

- <font size='5'>This would support the assumption that PPP is lower in low income economies as effective PPP fell along with GDP.</font>
::::

:::: column
```{r plot-Turkey, warning=FALSE, echo=FALSE, eval=TRUE}

bigmacTurkey <- big_mac_full_index %>%
  filter(name == "Turkey")


ggplot(data = bigmacTurkey, mapping = aes(x = date, y = USD_raw, colour = ifelse(USD_raw >0, "red", "blue"), group = 1)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0, colour = "darkolivegreen3",  size = 2) +
  theme(
    axis.text.x = element_text(angle = 45),
    legend.position = "none",
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "lightgrey"),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(colour = "hotpink")
  ) +
  labs(
    x = "Date",
    y = "Raw Index (dollars)",
    title = "Change in the Raw Index over time",
    subtitle = "Biannually for Turkey 2000-2021"
  )



```
::::
:::

## Switzerland 
::: columns
:::: column
- <font size='5'>The swiss franc has increased in value continually for a substantial time.</font>

- <font size='5'>It is seen as a safe haven for foreign exchange traders, who will usually exchange their money for francs during times of financial crisis.</font>

- <font size='5'>This has meant that the swiss have had high levels of PPP for a long period of time. Due to investor confidence in the franc. </font>

::::

:::: column
```{r plot-Switzerland, echo=FALSE, warning=FALSE}

bigmacSwitz <- big_mac_full_index %>%
  filter(name %in% c("Switzerland"))


ggplot(data = bigmacSwitz, mapping = aes(x = date, y = USD_raw, colour = ifelse(USD_raw >0, "red", "blue"), group = 1)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0, colour = "darkolivegreen3",  size = 2) +
  theme(
    axis.text.x = element_text(angle = 45),
    legend.position = "none",
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "lightgrey"),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(colour = "orchid4")
  ) +
  labs(
    x = "Date",
    y = "Raw Index (dollars)",
    title = "Change in the Raw Index over time",
    subtitle = "Biannually for Switzerland 2000-2021"
  )

```
::::
:::

## Hong Kong
::: columns
:::: column
- <font size='4'>Hong Kong is a high GDP country.</font>

- <font size='4'>It would be expected that the Hong Kong dollar would be overvalued due to Hong Kong's high level of GDP per capita and investor confidence. Meaning we would expect Hong Kong to have a high level of PPP</font>

- <font size='4'>However this is not the case as the HK$ is "pegged" against the American dollar. This means that the HK$ will change in value as the US$ changes in value. As the US$ is stable it will stay around the same level resulting in sustained low PPP for Hong Kong. </font>

::::

:::: column
```{r plot-HongKong, echo=FALSE, warning=FALSE}

bigmacHK <- big_mac_full_index %>%
  filter(name %in% c("Hong Kong"))


ggplot(data = bigmacHK, mapping = aes(x = date, y = USD_raw, colour = ifelse(USD_raw >0, "red", "blue"), group = 1)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0, colour = "darkolivegreen3",  size = 2) +
  theme(
    axis.text.x = element_text(angle = 45),
    legend.position = "none",
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "lightgrey"),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(colour = "forestgreen")
  ) +
  labs(
    x = "Date",
    y = "Raw Index (dollars)",
    title = "Change in the Raw Index over time",
    subtitle = "Biannually for Hong Kong 2000-2021"
  )

```
::::
:::

## Brazil
::: columns
:::: column
- <font size='4'>Brazil is a relatively low GDP country.</font>

- <font size='4'>Brazils PPP has varied from 2000 -2021 but in the period of 2006 to 2014 brazil had high levels of PPP.</font>

- <font size='4'>This was due to Brazil being one of the fastest growing economies in the world which huge levels of investor confidence. </font>

- <font size='4'>This fell after 2014 when brazil experienced one of its worst recorded recessions in the countries history due to falling comodity prices and an internal political struggle which shook investor confidence.</font>

::::

:::: column
```{r plot-Brazil, warning=FALSE, out.width="100%", echo=FALSE}

bigmacBrazil <- big_mac_full_index %>%
  filter(name %in% c("Brazil"))


ggplot(data = bigmacBrazil, mapping = aes(x = date, y = USD_raw, colour = ifelse(USD_raw >0, "red", "blue"), group = 1)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0, colour = "darkolivegreen3",  size = 2) +
  theme(
    axis.text.x = element_text(angle = 45),
    legend.position = "none",
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "lightgrey"),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(colour = "orange")
  ) +
  labs(
    x = "Date",
    y = "Raw Index (dollars)",
    title = "Change in the Raw Index over time",
    subtitle = "Biannually for Brazil 2000-2021"
  )

```
::::
:::


# Feeling adventurous?

You can learn more about the options available at the following pages:

- https://github.com/rstudio/revealjs

- https://bookdown.org/yihui/rmarkdown/revealjs.html

- http://www.jenniferbradham.org/post/reveal_js_markdown/

- https://revealjs.com/