---
title: "The Big-mac Index"
subtitle: <font size='6'>Purchasing Power Parity </font>
author: <font size='8'>The Code Ninjas <br> Tuilelaith, Dara, Kay Thari, Anya </font>
institute: <font size='10'>University of Edinburgh</font>
date: <font size = '6'>`r Sys.Date()`</font>
output:
  revealjs::revealjs_presentation:
    self_contained: false
editor_options: 
  chunk_output_type: console
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(rmarkdown)
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(revealjs)
library(gganimate)
library(gifski)
library(grid)
library(gridExtra)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
suppressWarnings(expr)
```

```{r load-data, include=FALSE}
big_mac_full_index = read.csv(file = "/cloud/project/data/big-mac-data-2021-07.1/output-data/big-mac-full-index.csv")
```

## <span style="color: turquoise;"> A statement of the overall goal / research question </span>  {data-background-image="img/Global-Economy-4.jpg"}
<span style="color: yellow;"> <font size='7'> Our research question is “What is Purchasing Power Parity (PPP) and how has it changed over the years across countries?” </font> </span>

# <span style="color: yellowgreen;"> What is Purchasing Power Parity? </span> {data-background-image="img/Blue gradient.jpg"}

<span style="color: snow;"> <font size='6'> PPP stands for Purchasing Power Parity.</font></span>

<span style="color: snow;"> <font size='6'> It is a measurement used in economics to compare the purchasing power of different currencies in common units. </font></span>

<span style="color: snow;"> <font size='6'> In theory, PPP should be equal to the exchange rate of these two countries, but in practice this is not the case.</font></span>

# <span style="color: royalblue;"> The Big-mac Index Data</span> {data-background-image="img/Burger2.jpg"}

<font size='6'> The dataset consists of 1,520 observations (rows) and 19 variables (columns).</font>

 <font size='6'>Each row represents a country at a specific time (year-month-date) while the columns represent crucial information related to the countries.</font>

<font size='6'>The data is accountable for the years 2000-2021. </font>

<font size='6'>The Big Mac prices are from McDonald’s, and the exchange rates are from Thomson Reuters. The GDP per person data is from the IMF World Economic Outlook reports.</font>

---

```{r tibble}
dim(big_mac_full_index)

ls(big_mac_full_index)

```

# <span style="color: seagreen;"> Why Big Macs?</span> {data-background-image="img/Yellow gradient.jpg"}

<span style="color: darkviolet;"> <font size='6'> The Big Mac Index is an accurate way of looking at PPP in the real world. </font></span>

<span style="color: darkviolet;"> <font size='6'> A Big Mac is going to be the same and hold the same intrinsic value no matter what country it is in, and so it is a good product to compare the currencies across countries.</font></span>

# <span style="color: royalblue;"> Analysis of PPP</span> {data-background-image="img/Burger2.jpg"}
<font size='6'>One of the common assumptions surrounding PPP is that a lower income country will have 
on average, lower prices compared to a high income country. This would be due to lower
labour costs and an overall inability to afford more costly goods in a low income economy.This will mean that a person in a low income economy will be able to buy less
in a high income economy than they could before. This would therefore give us the impression that PPP is lower in low income economies than in high income economies.</font>



# Plots - PPP by GDP per capita
::: columns
:::: column

- <font size='5'> The assumption that PPP is higher in high income countries is supported by this graph. Raw Index shows how overvalued/undervalued a countries currency is compared to the big mac price in each country.</font>

- <font size='5'> Effectively a country with a highly overvalued currency will have a high PPP and a country with a highly undervalued currency will have a low PPP.</font>


::::

:::: column
```{r plot-gdp/usdraw, echo = FALSE, warning=FALSE, out.height="100%", fig.height=5, message=FALSE}

bigmac2011 <- big_mac_full_index %>%
  filter(date >= as.Date("2011-01-01"), date <= as.Date("2011-12-31"))

ggplot(data = bigmac2011, mapping = aes(x = GDP_dollar, y = USD_raw )) +
  geom_point() +
  theme_minimal() +
  geom_smooth(method = "lm", se = FALSE, colour = "grey", linetype = "dashed",formula = y ~ x) +
  labs(
    x = "GDP per person (dollars)",
    y = "Raw Index (dollars)",
    title = "",
    subtitle = "For 2011",
    fill = "Countries"
  ) +
  theme(
    legend.position = "bottom"
  )



##line <- linear_reg() %>%
  ##set_engine("lm") %>%
  ##fit(USD_raw ~ GDP_dollar, data = bigmac2011) 


```
::::
:::

---

```{r plot-gdp/usdrawYRS, echo=FALSE, warning=FALSE, message=FALSE}


big_mac_full_index$year = str_sub(big_mac_full_index$date,1,4)

bigmac <- big_mac_full_index %>%
  filter(year %in% c("2012","2014","2016","2018","2020"))

bigmac %>%
  group_by(year) %>%
  ggplot(data = bigmac, mapping = aes(x = GDP_dollar, y = USD_raw, colour = name)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, colour = "grey", linetype = "dashed",formula = y ~ x) +
  labs(
    x = "GDP per person (dollars)",
    y = "Raw Index (dollars)",
    title = "",
    subtitle = "Through the decade"
  ) +
  theme(
    legend.position = "none",
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "lightgrey"),
    plot.title = element_text(face = "bold")
  )  +
  facet_grid(year ~ .)

```


## Turkey
::: columns
:::: column
- <font size='5'>High rates of GDP growth till 2007 with greater investor confidence.</font>

- <font size='5'>Post 2008, industry declined and led to higher levels of inflation and a reduction in imports, all of this led to the Lira fall in value up till 2021</font>

- <font size='5'>This would support the assumption that PPP is lower in low income economies as effective PPP fell along with GDP.</font>
::::

:::: column
```{r plot-Turkey, warning=FALSE, echo=FALSE, eval=TRUE}

bigmacTurkey <- big_mac_full_index %>%
  filter(name == "Turkey")


ggplot(data = bigmacTurkey, mapping = aes(x = date, y = USD_raw, colour = ifelse(USD_raw >0, "red", "blue"), group = 1)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0, colour = "darkolivegreen3",  size = 2) +
  theme(
    axis.text.x = element_text(angle = 45),
    legend.position = "none",
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "lightgrey"),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(colour = "hotpink")
  ) +
  labs(
    x = "Date",
    y = "Raw Index (dollars)",
    title = "Change in the Raw Index over time",
    subtitle = "Biannually for Turkey 2000-2021"
  )



```
::::
:::

## Switzerland 
::: columns
:::: column
* Test 1
* Test 2
::::

:::: column
```{r plot-Switzerland, echo=FALSE, warning=FALSE}

bigmacSwitz <- big_mac_full_index %>%
  filter(name %in% c("Switzerland"))


ggplot(data = bigmacSwitz, mapping = aes(x = date, y = USD_raw, colour = ifelse(USD_raw >0, "red", "blue"), group = 1)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0, colour = "darkolivegreen3",  size = 2) +
  theme(
    axis.text.x = element_text(angle = 45),
    legend.position = "none",
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "lightgrey"),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(colour = "orchid4")
  ) +
  labs(
    x = "Date",
    y = "Raw Index (dollars)",
    title = "Change in the Raw Index over time",
    subtitle = "Biannually for Switzerland 2000-2021"
  )

```
::::
:::

## Hong Kong
::: columns
:::: column
* Test 1
* Test 2
::::

:::: column
```{r plot-HongKong, echo=FALSE, warning=FALSE}

bigmacHK <- big_mac_full_index %>%
  filter(name %in% c("Hong Kong"))


ggplot(data = bigmacHK, mapping = aes(x = date, y = USD_raw, colour = ifelse(USD_raw >0, "red", "blue"), group = 1)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0, colour = "darkolivegreen3",  size = 2) +
  theme(
    axis.text.x = element_text(angle = 45),
    legend.position = "none",
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "lightgrey"),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(colour = "forestgreen")
  ) +
  labs(
    x = "Date",
    y = "Raw Index (dollars)",
    title = "Change in the Raw Index over time",
    subtitle = "Biannually for Hong Kong 2000-2021"
  )

```
::::
:::

## Brazil
::: columns
:::: column
* Test 1
* Test 2
::::

:::: column
```{r plot-Brazil, warning=FALSE, out.width="100%", echo=FALSE}

bigmacBrazil <- big_mac_full_index %>%
  filter(name %in% c("Brazil"))


ggplot(data = bigmacBrazil, mapping = aes(x = date, y = USD_raw, colour = ifelse(USD_raw >0, "red", "blue"), group = 1)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0, colour = "darkolivegreen3",  size = 2) +
  theme(
    axis.text.x = element_text(angle = 45),
    legend.position = "none",
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "lightgrey"),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(colour = "orange")
  ) +
  labs(
    x = "Date",
    y = "Raw Index (dollars)",
    title = "Change in the Raw Index over time",
    subtitle = "Biannually for Brazil 2000-2021"
  )

```
::::
:::


# Feeling adventurous?

You can learn more about the options available at the following pages:

- https://github.com/rstudio/revealjs

- https://bookdown.org/yihui/rmarkdown/revealjs.html

- http://www.jenniferbradham.org/post/reveal_js_markdown/

- https://revealjs.com/