---
title: "The Big-mac Index"
subtitle: <font size='6'>Purchasing Power Parity </font>
author: <font size='8'>The Code Ninjas <br> Tuilelaith, Dara, Kay Thari, Anya </font>
institute: <font size='10'>University of Edinburgh</font>
date: <font size = '6'>`r Sys.Date()`</font>
output:
  revealjs::revealjs_presentation:
    self_contained: false
editor_options: 
  chunk_output_type: console
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(rmarkdown)
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(revealjs)
library(gganimate)
library(gifski)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
suppressWarnings(expr)
```

```{r load-data, include=FALSE}
big_mac_full_index = read.csv(file = "/cloud/project/data/big-mac-data-2021-07.1/output-data/big-mac-full-index.csv")
```

## <span style="color: turquoise;"> A statement of the overall goal / research question </span>  {data-background-image="img/Global-Economy-4.jpg"}
<span style="color: yellow;"> <font size='7'> Our research question is “What is Purchasing Power Parity (PPP) and how has it changed over the years across countries?” </font> </span>

# <span style="color: yellowgreen;"> What is Purchasing Power Parity? </span> {data-background-image="img/Blue gradient.jpg"}

<span style="color: snow;"> <font size='6'> PPP stands for Purchasing Power Parity</font></span>

<span style="color: snow;"> <font size='6'> It is a measurement used in economics to compare the purchasing power of different currencies in common units. </font></span>

<span style="color: snow;"> <font size='6'> In theory, PPP should be equal to the exchange rate of these two countries, but in practice this is not the case.</font></span>

# <span style="color: royalblue;"> The Big-mac Index Data</span> {data-background-image="img/Burger2.jpg"}

<font size='6'> The dataset consists of 1,520 observations (rows) and 19 variables (columns).</font>

 <font size='6'>Each row represents a country at a specific time (year-month-date) while the columns represents crucial information related to the countries.</font>

<font size='6'>The data is accountable for the years 2000-2021. </font>

<font size='6'>The Big Mac prices are from McDonald’s, and the exchange rates are from Thomson Reuters. The GDP per person data is from the IMF World Economic Outlook reports.</font>

---

```{r tibble}
dim(big_mac_full_index)

ls(big_mac_full_index)

```


# <span style="color: seagreen;"> Why Big Macs?</span> {data-background-image="img/Yellow gradient.jpg"}

<span style="color: darkviolet;"> <font size='6'> The Big Mac Index is an accurate way of looking at PPP in the real world. </font></span>

<span style="color: darkviolet;"> <font size='6'> A Big Mac is going to be the same and hold the same intrinsic value no matter what country it is in, and so it is a good product to compare the currencies across countries.</font></span>

# Sub-questions and how we will answer them

It's possible to have a horizontal sliding transition as well as nest multiple slides to create a vertical stack.

When presenting, you use the left/right arrows to step through the top-level (horizontal) slides. When you arrive at a vertical stack you can optionally press the up/down arrows to view the vertical slides or skip past them by pressing the right arrow.

(click the down key now)

## reveal.js

- The presentation is created using the `reveal.js` package

- You can create a slide show broken up into sections by using the # and ## heading tags

- You can also create a new slide without a header using a horizontal rule (----)

----

- You can also add pauses between content on a slide using (". . .")

. . .

Like this.

...or using "::: fragment"

::: fragment

Like this.

:::

# Layouts

You can use plain text

- or bullet points

<div>
  <div style="float: left; width: 50%;">or text in two columns</div>
  <div style="float: left; width: 50%;">like this</div>
</div>

because you can combine HTML and Markdown syntax.

<p style="font-size: 8pt; position: absolute; bottom: -5em; width: 100%; text-align: center;">*You can also add footnotes</p>

# Code

```{r boring-regression}
# a boring regression
model <- lm(dist ~ speed, data = cars)
tidy(model)
glance(model)
```

# Plots


```{r plot-gdp/usdraw, echo = FALSE, warning=FALSE, out.height="100%", fig.height=5}

bigmac2011 <- big_mac_full_index %>%
  filter(date >= as.Date("2011-01-01"), date <= as.Date("2011-12-31"))

ggplot(data = bigmac2011, mapping = aes(x = GDP_dollar, y = USD_raw )) +
  geom_point() +
  theme_minimal() +
  geom_smooth(method = "lm", se = FALSE, colour = "grey", linetype = "dashed",formula = y ~ x) +
  labs(
    x = "GDP per person (dollars)",
    y = "Raw Index (dollars)",
    title = "",
    subtitle = "For 2011",
    fill = "Countries"
  ) +
  theme(
    legend.position = "bottom"
  )

linear_reg() %>%
  set_engine("lm") %>%
  fit(USD_raw ~ GDP_dollar, data = bigmac2011) %>%
  tidy()
  

```

## Animated

```{r plot-gdp/usdrawYRS, animation.hook='gifski'}

big_mac_full_index$year = str_sub(big_mac_full_index$date,1,4)

bigmac <- big_mac_full_index %>%
  filter(year %in% c("2011","2012","2013","2014","2015","2016","2017","2018","2019","2020","2021"))
  
anim <- bigmac %>%  
  ggplot(data = bigmac, mapping = aes(x = GDP_dollar, y = USD_raw, colour = name)) +
  geom_point() +
  transition_time(readr::parse_number(year)) +
  labs(
    x = "GDP per person (dollars)",
    y = "Raw Index (dollars)", 
    title = "",
    subtitle = "Through Years: {as.integer(frame_time)}",
    fill = "Countries"
  ) +
  theme(
    legend.position = "none",
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "lightgrey"),
    plot.title = element_text(face = "bold")
  )  

#animate(anim, fps = 2, renderer = gifski_renderer())


```


## Turkey
```{r plot-Turkey, warning=FALSE, echo=FALSE, eval=TRUE}

bigmacTurkey <- bigmac %>%
  filter(name %in% c("Turkey"))


ggplot(data = bigmacTurkey, mapping = aes(x = date, y = USD_raw, colour = ifelse(USD_raw >0, "red", "blue"), group = 1)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0, colour = "darkolivegreen3",  size = 2) +
  theme(
    axis.text.x = element_text(angle = 45),
    legend.position = "none",
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "lightgrey"),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(colour = "hotpink")
  ) +
  labs(
    x = "Date",
    y = "Raw Index (dollars)",
    title = "Change in the Raw Index over time",
    subtitle = "Biannually for Turkey 2011-2021"
  )



```

## Switzerland
```{r plot-Switzerland, echo=FALSE, warning=FALSE}

bigmacSwitz <- bigmac %>%
  filter(name %in% c("Switzerland"))


ggplot(data = bigmacSwitz, mapping = aes(x = date, y = USD_raw, colour = ifelse(USD_raw >0, "red", "blue"), group = 1)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0, colour = "darkolivegreen3",  size = 2) +
  theme(
    axis.text.x = element_text(angle = 45),
    legend.position = "none",
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "lightgrey"),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(colour = "orchid4")
  ) +
  labs(
    x = "Date",
    y = "Raw Index (dollars)",
    title = "Change in the Raw Index over time",
    subtitle = "Biannually for Switzerland 2011-2021"
  )

```

## Hong Kong
```{r plot-HongKong, echo=FALSE, warning=FALSE}

bigmacHK <- bigmac %>%
  filter(name %in% c("Hong Kong"))


ggplot(data = bigmacHK, mapping = aes(x = date, y = USD_raw, colour = ifelse(USD_raw >0, "red", "blue"), group = 1)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0, colour = "darkolivegreen3",  size = 2) +
  theme(
    axis.text.x = element_text(angle = 45),
    legend.position = "none",
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "lightgrey"),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(colour = "forestgreen")
  ) +
  labs(
    x = "Date",
    y = "Raw Index (dollars)",
    title = "Change in the Raw Index over time",
    subtitle = "Biannually for Hong Kong 2011-2021"
  )

```

## Brazil
```{r plot-Brazil, warning=FALSE, out.width="100%", echo=FALSE}

bigmacBrazil <- bigmac %>%
  filter(name %in% c("Brazil"))


ggplot(data = bigmacBrazil, mapping = aes(x = date, y = USD_raw, colour = ifelse(USD_raw >0, "red", "blue"), group = 1)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0, colour = "darkolivegreen3",  size = 2) +
  theme(
    axis.text.x = element_text(angle = 45),
    legend.position = "none",
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "lightgrey"),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(colour = "orange")
  ) +
  labs(
    x = "Date",
    y = "Raw Index (dollars)",
    title = "Change in the Raw Index over time",
    subtitle = "Biannually for Brazil 2011-2021"
  )

```

```{r recode-species, echo = FALSE}
# In this chunk I'm doing a bunch of analysis that I don't want to present 
# in my slides. But I need the resulting data frame for a plot I want to present.
iris_modified <- iris %>%
  mutate(Species = fct_other(Species, keep = "setosa"))
```

```{r plot-iris, echo = FALSE}
# Code hidden with echo = FALSE
# Uses modified iris dataset from previous chunk
# Play around with height and width until you're happy with the look
ggplot(data = iris_modified, mapping = aes(x = Sepal.Width, y = Sepal.Length, color = Species)) +
  geom_point() + 
  theme_minimal() # theme options: https://ggplot2.tidyverse.org/reference/ggtheme.html
```

## Plot and text

::: columns
:::: column
* Some text
* goes here
::::

:::: column
```{r warning=FALSE, out.width="100%", fig.width=4, echo=FALSE}
# see how I changed out.width and fig.width from defaults
# to make the figure bigger
ggplot(penguins, aes(x = bill_length_mm, y = species, color = species)) +
  geom_boxplot() +
  theme_minimal()
```
::::
:::

## Different widths

::: columns
:::: {.column width="40%"}
* Some text
* goes here
::::

:::: {.column width="60%"}
```{r warning=FALSE, out.width="100%", fig.width=4, echo=FALSE}
# see how I changed out.width and fig.width from defaults
# to make the figure bigger
ggplot(penguins, aes(x = bill_length_mm, y = species, color = species)) +
  geom_boxplot() +
  theme_minimal()
```
::::
:::

# Tables

If you want to generate a table, make sure it is in the HTML format (instead of Markdown or other formats), e.g.,

```{r iris-table, echo = FALSE}
kable(head(iris), format = "html")
```

# Images

```{r castle, echo = FALSE, out.width = "60%", fig.align = "center", fig.cap = "Image credit: Photo by Jörg Angeli on Unsplash."}
include_graphics("https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80")
```

Or you can also include a full page image. See next slide.

## {data-background-image="https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80"}

# Math Expressions

You can write LaTeX math expressions inside a pair of dollar signs (\$...\$), e.g. $\alpha+\beta$. You can use the display style with double dollar signs:

$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$

# Feeling adventurous?

You can learn more about the options available at the following pages:

- https://github.com/rstudio/revealjs

- https://bookdown.org/yihui/rmarkdown/revealjs.html

- http://www.jenniferbradham.org/post/reveal_js_markdown/

- https://revealjs.com/